---
title: "K-means-practice"
author: Izolda Fetko
date: 3/24/2019
output: html_document
---

The following practive was retrieved from https://bookdown.org/rdpeng/exdata/k-means-clustering.html

First we need to import necessary libraries.

```{r message=FALSE, warning=FALSE}
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
```

The K-means approach, like many clustering methods, is highly algorithmic (can't be summarized in a formula) and is iterative. 
The outline of the algorithm is:

- Fix the number of clusters at some integer greater than or equal to 2

- Start with the "centroids" of each cluster; initially you might just pick a random set of points as the centroids

- Assign points to their closest centroid; cluster membership corresponds to the centroid assignment

- Reclaculate centroid positions and repeat.

To illustrate the example we are working with the random dataset.

```{r}
set.seed(1234)
x <- rnorm(12, mean = rep(1:3, each = 4), sd = 0.2)
y <- rnorm(12, mean = rep(c(1, 2, 1), each = 4), sd = 0.2)
plot(x, y, col = "blue", pch = 19, cex = 2)
text(x + 0.05, y + 0.05, labels = as.character(1:12))

```

The first thing K-means has to do is assign an initial set of centroids.
The next stage in the algorithm assigns every point in the dataset to the closest centroid.
The next stage is the re-calculate the centroids based on the new cluster assignments of the data points. 
Once the full cycle of the althorithm is completed, re-assign points to their (new) closest cluster centroid.
We can also update the centroid positions one more time based on the re-assigned points.

### Kmeans() function

The kmeans() function in R implements the K-means algorithm and can be found in the stats package, which comes with R.

Two key parameters thathavet to be specified are x, which is a matrix or data frame of data, and centers which is either an integer indicating the number of clusters or a matrix indicating the locations of the initial cluster centroids. The data should be organized so that each row is an observation and each column is a variable or feature of that observation.

```{r}
dataFrame <- data.frame(x, y)
kmeansObj <- kmeans(dataFrame, centers = 3)
names(kmeansObj)

```

You can see which cluster each data point got assigned to by looking at the cluster element of the list returned by the kmeans() function.

```{r}
kmeansObj$cluster
```

Once we have the k-means algorithm solutions, we can build heat maps from them.

```{r}
set.seed(1234)
dataMatrix <- as.matrix(dataFrame)[sample(1:12), ]
kmeansObj <- kmeans(dataMatrix, centers = 3)
```

```{r}
par(mfrow = c(1, 2))
image(t(dataMatrix)[, nrow(dataMatrix):1], yaxt = "n", main = "Original Data")
image(t(dataMatrix)[, order(kmeansObj$cluster)], yaxt = "n", main = "Clustered Data")
```

