---
title: "K-means-practice"
author: Izolda Fetko
date: 3/24/2019
output: html_document
---

The following practive was retrieved from Canvas.

```{r message=FALSE, warning=FALSE}
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
```
Adapted from UC Business Analytics R Programing Guide

To be able to conduct a K-means clustering analysis, first we need to prepare the dataset - remove the missing values and scale the dataset.

## Data Preparation

Dataset - 1973

```{r USAarests}
df <- USArrests
```

Remove missing values
```{r}
df <- na.omit(df)
```
Scaling
```{r}
df <- scale(df)

```
Now we run the K-Means algorithm

## K-Means
```{r}
k2 <- kmeans(df, centers = 2)  # centers  = K
str(k2)
```
Output 
- cluster: A vector of integers (from 1:k) indicating the cluster to which each point is allocated.
- centers: A matrix of cluster centers.
- totss: The total sum of squares.
- withinss: Vector of within-cluster sum of squares, one component per cluster.
- tot.withinss: Total within-cluster sum of squares, i.e. sum(withinss).
- betweenss: The between-cluster sum of squares, i.e. $totss-tot.withinss$.
- size: The number of points in each cluster.

```{r}
# function to compute total within-cluster sum of square 
wss <- function(k) {
  kmeans(df, k, nstart = 10 )$tot.withinss
}

# Compute and plot wss for k = 1 to k = 15
k.values <- 1:15

# extract wss for 2-15 clusters
wss_values <- map_dbl(k.values, wss)

plot(k.values, wss_values,
       type="b", pch = 19, frame = FALSE, 
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")
```
We can also use the Elbom method and nicely visualize the clusters.

### Elbow Method
```{r}
final <- kmeans(df, 4, nstart = 25)
fviz_cluster(final, data = df)
```