---
title: "Tidy data"
author: "Izolda Fetko"
date: "2/4/2019"
output:
  html_document:
    df_print: paged
---

### Installing all packages needed


```{r echo=FALSE, warning=FALSE, message = FALSE}
library(dplyr)
library(htmlwidgets)
library(tidyverse)
library(readr)
library(tidyr)
library(ggplot2)
```

#### Look at data in table1
```{r}
table1
```

#### Compute rate per 10,000

```{r}
table1 %>% 
  mutate(rate = cases / population * 10000)
```

#### Visualize changes over time

```{r}
library(ggplot2)
ggplot(table1, aes(year, cases)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
```

#### Exercise 1

**Using prose, describe how the variables and observations are organised in each of the sample tables.**

Table1 - rows represent (country, year) combination. The columns are cases and population.

Table2 - rows represent (country, year, variable) combination. The column *count* contains the values of variables for cases and population in separate rows.

Table3- rows represent (country, year) combination. The column *rate* provides the values of both cases and population combined in a string.

Table4 - is split into two tables, one table for each variable. Table4a contains cases and table4b contains populations.


**Compute the rate for table2, and table4a + table4b. You will need to perform four operations:**

**Extract the number of TB cases per country per year.**
**Extract the matching population per country per year.**
**Divide cases by population, and multiply by 10000.**
**Store back in the appropriate place.**


```{r}
t2_cases <- filter(table2, type == "cases") %>%
  dplyr::rename(cases = count) %>%
  arrange(country, year)
t2_population <- filter(table2, type == "population") %>%
  dplyr::rename(population = count) %>%
  arrange(country, year)
```


```{r}
t2_cases_per_cap <- tibble(
  year = t2_cases$year,
  country = t2_cases$country,
  cases = t2_cases$cases,
  population = t2_population$population
) %>%
  mutate(cases_per_cap = (cases / population) * 10000) %>%
  select(country, year, cases_per_cap)
```

```{r}
t2_cases_per_cap <- t2_cases_per_cap %>%
  mutate(type = "cases_per_cap") %>%
  dplyr::rename(count = cases_per_cap)
```

```{r}
bind_rows(table2, t2_cases_per_cap) %>%
  arrange(country, year, type, count)
```

```{r}
table4c <-
  tibble(
    country = table4a$country,
    `1999` = table4a[["1999"]] / table4b[["1999"]] * 10000,
    `2000` = table4a[["2000"]] / table4b[["2000"]] * 10000
  )
table4c
```
**Which representation is easiest to work with? Which is hardest? Why?**

I didn't find any of the tables easy to work with. Just because the tables are not structured well for example where the columns are not combined and we have only 4 columns.


#### Spreading and gathering exercise

**Why are gather() and spread() not perfectly symmetrical?**
**Carefully consider the following example:**

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`)
```

(Hint: look at the variable types and think about column names.)

**Both spread() and gather() have a convert argument. What does it do?**

They are not symmetrical because the column type information is lost.

**Why does this code fail?**

table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")

It fails because the column names `1999` and `2000` are not non-syntactic variable names

**Why does spreading this tibble fail? How could you add a new column to fix the problem?**

```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
```
To spread we would use: spread(people, key, value)

It fails because the columns key and name do not uniquely identify rows.


**Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?**

```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)

```

```{r}
preg_tidy <- preg %>%
  gather(male, female, key = "sex", value = "count")
preg_tidy
```
```{r}
preg_tidy2 <- preg %>%
  gather(male, female, key = "sex", value = "count", na.rm = TRUE)
preg_tidy2
```
To clean this data we need to remove the males, essentially assign "no" to the pregnant category for all males.

#### Separating and Uniting Exercise

**What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.**

```{r}

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
```

**Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?**
This argument allows you to remove orginal columns. So we would set it to FALSE if we wanted to check the original columns.

**Compare and contrast separate() and extract(). Why are there three variations of separation (by position, by separator, and with groups), but only one unite?**

Because we can separate items in different ways.


#### Missing Values Exercise

**Compare and contrast the fill arguments to spread() and complete().**

In `spread()` this argment  will replace all missing values regardless of columns with the same value and in the complete() it will receive a list where each slot is the missing value for each column.

**What does the direction argument to fill() do?**

It determines whether `NA` values should be replaced by the previousor the next non-missing value.



