---
title: "Managing Data Frames with the dplyr package"
author: "Izolda Fetko"
date: "2/4/2019"
output:
  html_document:
    df_print: paged
---

### Installing the dplyr package


```{r echo=FALSE, warning=FALSE, message = FALSE}
library(dplyr)
library(htmlwidgets)
library(tidyverse)
library(readr)
library(tidyr)

```

####Import Chicago data
```{r}
chicago <- readRDS("C:/Users/Justin/Downloads/chicago_data/chicago.rds")
```

#### SELECT()

To see some basic characteristics of the dataset, we can use dim() and str()

```{r}

dim(chicago)
str(chicago)
```

To take thefirst three columns only, we can do this:

```{r}
names(chicago)[1:3]
```

Also we can subset the data using select()

```{r}
subset <- select(chicago, city:dptp)
head(subset)
```

The select() function also allows a special syntax that allows you to specify variable names based on patterns. So, for example, if you wanted to keep every variable that ends with a "2", we could do

```{r}
subset <- select(chicago, ends_with("2"))
str(subset)
```

#### FILTER()

Example of how to filter the chicago data with mean greater than 30

```{r}
chic.f <- filter(chicago, pm25tmean2 > 30)
str(chic.f)
```

We can also use summary to see the summary statistics

```{r}
summary(chic.f$pm25tmean2)
```

We can place an arbitrarily complex logical sequence inside of filter(), so we could for example extract the rows where PM2.5 is greater than 30 and temperature is greater than 80 degrees Fahrenheit.

```{r}
chic.f <- filter(chicago, pm25tmean2 > 30 & tmpd > 80)
select(chic.f, date, tmpd, pm25tmean2)
   
```

#### ARRANGE()

We can use this function to rearrange rows of the dataframe according to some variables or columns.

```{r}

chicago <- arrange(chicago, date)

head(select(chicago, date, pm25tmean2), 3)
tail(select(chicago, date, pm25tmean2), 3)

```

We can sort columns in descending order and check the data by using the head function

```{r}

chicago <- arrange(chicago, desc(date))
head(select(chicago, date, pm25tmean2), 3)
```

#### MUTATE()

This function exists to compute transformations of variables in a data frame.

```{r}
chicago <- mutate(chicago, pm25detrend = pm25tmean2 - mean(pm25tmean2, na.rm = TRUE))
head(chicago)
```

#### GROUP_BY()
The group_by() function is used to generate summary statistics from the data frame within strata defined by a variable.

First we can create the year variable

```{r}
chicago <- mutate(chicago, year = as.POSIXlt(date)$year + 1900)
```

And now we can create another dataframe  and group data by year

```{r}
years <- group_by(chicago, year)
```

We do all this so that we can compute summary statistics by year.

```{r}
summarize(years, pm25tmean2 = mean(pm25tmean2, na.rm = TRUE), 
           o3 = max(o3tmean2, na.rm = TRUE), 
           no2 = median(no2tmean2, na.rm = TRUE))
```

#### Pipline Operator %>%

The pipeline operater %>% is very handy for stringing together multiple dplyr functions in a sequence of operations.

For example, we can combine mutliple R statements into a function

```{r}
mutate(chicago, month = as.POSIXlt(date)$mon + 1) %>% 
         group_by(month) %>% 
         summarize(pm25tmean2 = mean(pm25tmean2, na.rm = TRUE), 
                   o3 = max(o3tmean2, na.rm = TRUE), 
                   no2 = median(no2tmean2, na.rm = TRUE))
```