---
title: "Tutorial Compiled by Joan Luciano I-590 Spring 2018"
author: "Izolda Fetko"
date: "4/1/2019"
output:
  html_document:
    df_print: paged
---

The following exercise was retrieved from Module 12 in Canvas



### Import, Data Set-up and Exploration
Please note that I am using my husbands old computer for school purposes. His name is Justin.

```{r}
auto <- read.table("C:/Users/Justin/Dropbox/IU/INFO I590 APPLIED DATA SCIENCE/Exercises/auto-mpg.data",
                   col.names = c("MPG","cylinders","displacement","horsepower",
                                  "weight","acceleration","year","origin","name"))
str(auto)
summary(auto)
auto$horsepower <- as.integer(auto$horsepower)
hist(auto$horsepower)
auto$fuelEfficient <- as.logical(auto$MPG>30)
table(auto$fuelEfficient)
pie(table(auto$fuelEfficient), main="Fuel Efficient or Not")

```

Now we are splitting the dataset into test and train
```{r}
smp_size <- floor(.8 * nrow(auto))
set.seed(seed=2018)
ti <- sample(seq_len(nrow(auto)), size=smp_size) 
train.auto <- auto[ti,]
test.auto <- auto[-ti,]
str(train.auto)
str(test.auto)

```

Linear Regression
```{r}
mylm <- lm(MPG~weight+year+horsepower, train.auto)
res <- predict.lm(mylm, test.auto)
summary(abs(res - test.auto$MPG))
```

K-means clustering
```{r}
mykm <- kmeans(auto[c("weight","year")], 2)
plot(auto[c("weight","year")], col=mykm$cluster, pch=as.integer(auto$fuelEfficient))
mykm <- data.frame(auto, mykm$cluster)
with(mykm, table(fuelEfficient, mykm.cluster))
```

Need to import library rpart  and carry out the classification tree

```{r}
library("rpart")
ctree <- rpart(fuelEfficient ~ cylinders+horsepower+weight+acceleration+year,
              data = train.auto,
              method = 'class')
ctree.test <- data.frame(test.auto,predict(ctree,test.auto))
summary(ctree.test[which(ctree.test$fuelEfficient==T), c(10,11,12)])
summary(ctree.test[which(ctree.test$fuelEfficient==F), c(10,11,12)])
```

And finally carry out the regression tree analysis

```{r}
rtree <- rpart(MPG ~ cylinders+horsepower+weight+acceleration+year,
              data = train.auto,
              method = 'anova')
rtree.test <- data.frame(test.auto, predict(rtree,test.auto))
summary(abs(rtree.test[1]-rtree.test[11]))
cor(rtree.test[1], rtree.test[11], method = "pearson")
```
