---
title: "Data Visualization Practice"
author: "Izolda Fetko"
date: "2/20/2019"
output:
  html_document:
    df_print: paged
---

The following exercise was retrieved from https://r4ds.had.co.nz/data-visualisation.html


### Installing the dplyr package


```{r echo=FALSE, warning=FALSE, message = FALSE}
library(dplyr)
library(htmlwidgets)
library(tidyverse)
library(readr)
library(tidyr)

```

view mpg data
```{r}
mpg
```

How many rows are in mtcars? How many columns? 32 rows and 11 columns

```{r}
nrow(mtcars)
ncol(mtcars)
```

What does the drv variable describe? Read the help for ?mpg to find out. 
The drv variable indicates the drive of the vehicle - front, rear, or 4 wheel drive.

Make a scatterplot of hwy vs cyl.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = hwy, y = cyl))
```

What happens if you make a scatterplot of class vs drv. Why is the plot not useful? 
Because both variables are categorical, the points overlap. The best practice is to have on categorical and
one numerical variable

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = class, y = drv))
```

Plot mpg data displ on x axis and hwy on y axis

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```


Assign color to each class of the car

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

Instead of mapping clas to color, now we map it to size of points

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

To control the transparency of datapoints, we assign class to the alpha (aesthetic) different point transparency

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

Different point shapes - returns warning of high number of shapes

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

Setting the geom aesthetics manually - datapoints to be blue

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

Qhy aren't these datapoints blue? Because the color argument was set within the wrong method,
aes(), and was supposed to be set in geom_point()
y = hwy

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

Which variables in mpg are categorical? Which variables are continuous? The categorical values are:
manufacturer, model, trans, drv, fl, class, The continuous are: displ, cyl, cty, hwy

Map a continuous variable to color, size, and shape. How do these aesthetics behave 
differently for categorical vs. continuous variables?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cty))
```


Shape - We get an error: A continuous variable can not be mapped to shape


ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = cty))

What happens if you map the same variable to multiple aesthetics? all of them will be implemented
and the number of legends will be equal to number of aesthetics

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = cty))
```

Does the stroke aesthetic do? What shapes does it work with? It affects the border thickness 
for shapes that can take on different colors both inside and outside

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), stroke = 3, shape = 21)
```

What happens if you map an aesthetic to something other than a variable name, like  
aes(colour = displ < 5)? The code will be executed with a temporary variable and hold the
resutls. 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```

#### facets

What happens if you facet on a continuous variable? A separate facet will be drawn for each unique
value of that variable.

What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?
Empty cells mean there is no data. in the data that have that unique combination of values.
the variable name in the facet_wrap() is actually the forumla

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```

What plots does the following code make? What does . do? Dot is the placeholder for a variable

```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

Take the first faceted plot in this section. What are the advantages to using faceting instead 
of the colour aesthetic? What are the disadvantages? How might the balance change if you had a 
larger dataset? Facets help split the data into multiple groups/grids which helps to better visualize it.
The disadvantage is that splitting minimizes the visibility of the overall data relationships across
facets. To balance the change we use color in the following way - use it for smaller datasets because 
in larger datasets the points may overlap and minimize visiblity.


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the 
individual panels? Why doesn't facet_grid() have nrow and ncol arguments? 

nrow determines row count for faceted plots 
ncol same as nrow just for columns
as.table sets the starting facet to begin filling the plot, 
dir sets the  filling direction

When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?
it make the plot more visible because it is extended

####Geometric objects

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))

ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))

ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```

Displaying geoms in the same plot

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

Geom with different mappings

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()
```

Specifying data for each layer

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
 Line chart - geom_line()
 Boxplot - geom_boxplot()
 Histogram - geom_histogram()
 Area chart - geom_area()

Run this code in your head and predict what the output will look like. 
Then, run the code in R and check your predictions.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter? It removes the legend from the plot.

What does the se argument to geom_smooth() do? It allows to draw a conf interval.

Will these two graphs look different? Why/why not? They look the same. The groupings are
the same.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

Recreate the R code necessary to generate the following graphs.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(aes(group = drv), se = FALSE) +
  geom_point()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(aes(color = drv)) + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(aes(color = drv)) +
  geom_smooth(aes(linetype = drv), se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(size = 4, colour = "white") + 
  geom_point(aes(colour = drv))
```

####Statistical Transformations

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

The same graph can be created using stat_count()

To draw greater attention to the statistical transformation use stat_summary()

```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

What is the default geom associated with stat_summary()? How could you rewrite 
the previous plot to use that geom function instead of the stat function?

The default geom is geom_pointrange()

```{r}
ggplot(data = diamonds) +
  geom_pointrange(mapping = aes(x = cut, y = depth),
                  stat = "summary",
                  fun.ymin = min,
                  fun.ymax = max,
                  fun.y = median)
```
  
What does geom_col() do? How is it different to geom_bar()? they are the same in cases where
geom_bar includes stat = "identity".
  
What variables does stat_smooth() compute? What parameters control its behaviour? 

It calculates four variables: y - predicted value, #ymin - lower pointwise confidence interval around the mean,
ymax - upper pointwise confidence interval around the mean, #se - standard error. it controls the method that
plots the confidence interval
  

In our proportion bar chart, we need to set group = 1. Why? In other words what is the 
problem with these two graphs? 

If we don't do that, the proportions will be calculated based on the entire dataset, not just the subset
  
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```

####Position adjustments

We can color the bar chart

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
```

We can use cut and clarity also

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

What is the problem with this plot? How could you improve it? It is apparent that some
datapoints are invisible due to overlap. To fix this we can jitter the graph.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```

What parameters to geom_jitter() control the amount of jittering?
weight and height

Compare and contrast geom_jitter() with geom_count(). The first one counts the observations by location 
which are use to map the count to points area. The number of visible points is equal to geom_points()


What's the default position adjustment for geom_boxplot()? Create a visualisation of 
the mpg dataset that demonstrates it.

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy, color = drv)) + 
  geom_boxplot(position = "dodge")
```

####Coordinate systems


coord_flip()

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```

Turn a stacked bar chart into a pie chart using coord_polar().

```{r}

ggplot(data = mpg, mapping = aes(x = factor(1), fill = class)) +
  geom_bar(width = 1) +
  coord_polar(theta = "y")
```

What does labs() do? Read the documentation. labs() adds labels to the graph. 

What's the difference between coord_quickmap() and coord_map()?

coord_map() projects a portion of the earth (a three-dimensional object) onto a flat 
(two-dimensional) plane. coord_map() does not preserve straight lines and therefore 
is computationally intensive;  coord_quickmap() preserves straight lines and is therefore 
faster to draw (though less accurate).

What does the plot below tell you about the relationship between city and highway mpg? 
Why is  coord_fixed() important? What does geom_abline() do?

Overall all cars have a better mileage on highway that in the city. coord_fixed() helps draw equal 
intervals on both axes which makes them nicely comparavle. geom_abline() draws a line with intercept 0 
and slope 1 which again helps us make a better comparison between the two.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```