---
title: "Data Analysis Examples - Good Diagnostics"
author: "Izolda Fetko"
date: "3/19/2019"
output:
  html_document:
    df_print: paged
---

The following exercise was retrieved from http://daviddalpiaz.github.io/appliedstats/model-diagnostics.html#data-analysis-examples

Note: Was unable to complete Suspect Diagnostics sections due to the devtools package issues.


### Install packages we usually use in  exercises


```{r echo=FALSE, warning=FALSE, message = FALSE}
library(dplyr)
library(htmlwidgets)
library(tidyverse)
library(readr)
library(tidyr)

```

First we need to import the mtcars dataset

```{r}
mtcars
```

Now we need to fit the linear regression model

```{r}
mpg_hp_add = lm(mpg ~ hp + am, data = mtcars)

```

We plot to see the the relationship between the fitted versus residuals.
```{r}
plot(fitted(mpg_hp_add), resid(mpg_hp_add), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residual",
     main = "mtcars: Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)
```
```{r}
qqnorm(resid(mpg_hp_add), col = "darkgrey")
qqline(resid(mpg_hp_add), col = "dodgerblue", lwd = 2)
```
By plotting the qq plot, we see that the fitted vs residual still looks good.

```{r}
sum(hatvalues(mpg_hp_add) > 2 * mean(hatvalues(mpg_hp_add)))
```
This shows us that there are two points of large leverage.
To check if there is a point with a large residual we do the following:
```{r}
cd_mpg_hp_add = cooks.distance(mpg_hp_add)
sum(cd_mpg_hp_add > 4 / length(cd_mpg_hp_add))
```
```{r}
large_cd_mpg = cd_mpg_hp_add > 4 / length(cd_mpg_hp_add)
cd_mpg_hp_add[large_cd_mpg]
```
These two cars represent two influental points.

Checking the coefficients.

```{r}
coef(mpg_hp_add)

mpg_hp_add_fix = lm(mpg ~ hp + am,
                    data = mtcars,
                    subset = cd_mpg_hp_add <= 4 / length(cd_mpg_hp_add))
coef(mpg_hp_add_fix)
```

We can also plot this

```{r}
par(mfrow = c(2, 2))
plot(mpg_hp_add)
```

For the suspect diagnostics, we can analyze the autompg dataset.
Before we start using the dataset, we need to install appropriate packatged and libraries.



