---
title: "Exploratory Data Analysis Checklist"
author: "Izolda Fetko"
date: "2/25/2019"
output:
  html_document:
    df_print: paged
---


1. Formulate your question

2. Read in your data

3. Check the packaging

4. Run str()

5. Look at the top and the bottom of your data

6. Check your "n"s

7. Validate with at least one external data source

8. Try the easy solution first

9.Challenge your solution

10. Follow up

Installing all necessary packages


```{r}
library(dplyr)
library(htmlwidgets)
library(tidyverse)
library(readr)
library(tidyr)
library(readr)

```

Read in the data

```{r}
ozone <- read_csv("C:/Users/Justin/Desktop/USEPAdata 017.csv")

```

```{r}
names(ozone) <- make.names(names(ozone))
```

Checking the number of rows and columns

```{r}
nrow(ozone)
ncol(ozone)
```

We can check the structure of the dataset by running the str(ozone) command

```{r}
str(ozone)
```

We can look at the top and bottom rows of the dataset to get familiar with it

```{r}
head(ozone)
tail(ozone)
```
####Select a numeric column and practice mean(), median(), sd(), var(), IQR(), range(), diff(range()) 

For this purpose, I am choosing the observation count column

##### mean

```{r}
mean(ozone$Observation.Count)
```

##### standard deviation

```{r}
sd(ozone$Observation.Count)

```

##### median

```{r}
median(ozone$Observation.Count)
```

##### variance
```{r}
var(ozone$Observation.Count)
```

##### inter-quartile range
```{r}
IQR(ozone$Observation.Count)
```

##### range

```{r}
range(ozone$Observation.Count)
```
##### diff range

```{r}
diff(range(ozone$Observation.Count))
```



